An Adaptive Algorithm Employing Continuous Linear Functionals

### 2019-06-03 ReplyToReferee2019-Apr.tex

\documentclass[letter]{article}
\usepackage{bbm,mathtools,array,longtable,booktabs,graphicx,color,enumitem}

\definecolor{orange}{rgb}{1.0,0.3,0.0}
\definecolor{violet}{rgb}{0.75,0,1}
\newcommand{\frednote}[1]{  {\textcolor{red}  {\mbox{**Fred:} #1}}}
\newcommand{\yuhannote}[1]{ {\textcolor{violet}  {\mbox{**Yuhan:} #1}}}
\newcommand{\tonynote}[1]{ {\textcolor{orange}  {\mbox{**Tony:} #1}}}

\textwidth 6.5in
\textheight 9in
\hoffset -1in
\voffset -1in
\input MACROS.tex
\begin{document}
\title{Response to the Referees}
\maketitle

\noindent We would like to thank you for your careful reviews of our manuscript and your helpful comments.

\noindent Below, we respond to the points made in your reports.

@@

\subsection*{Report beginning ``The submitted paper \ldots''}

While it is less traditional in numerical analysis to assume that a function belongs to a cone than to a ball, we assert that there are advantages.  We think of the difficulty of a problem not as the computational cost but as the ability for an algorithm to correctly determine the computational cost. Thus, problems inside our cones are easy, but they could be costly depending on the norm of the input function.

Our example in Figure 1 illustrates what an easy problem is.  When the Fourier coefficients of the input decay steadily, we can easily construct an adaptive algorithm that approximates the solution well, regardless of how large the norm of the input function is.  On the other hand, when the decay is not steady, this is a difficult problem because the adaptive algorithm may be fooled unless the initial sample size is quite large.  Our cones are non-convex, which is consistent with what is known about when adaptive algorithms might be effective \cite{Bak71}.

\begin{enumerate}
    \item The first sentence is Subsection 1.1 has been rephrased.
    \item The sufficient and necessary conditions of the existence of singular value decomposition has been added  to  the last paragraph of Subsection 1.1
    \item \textit{linear functionals are used} changed to 
    \textit{linear functionals that are used}
    \item Inserted \textit{the} before the algorithm
    \item The norm for $\calF$ has been identified.
\end{enumerate}

@@

\subsection*{Report beginning ``This paper proposes \ldots''}

\begin{enumerate}[labelwidth = 10ex]
    \item[$1^{\text{st}}$ \bullet] Replaced
    $\sum\limits_{i =1 }^n$ by $\sum\limits_{i \in \mathbb{N} }$
    \item[$2^{\text{nd}}$ \bullet] Added one paragraph to end of \S 1.1 to foreshadow Example 1 and the numerical example.
    \item[$3^{\text{rd}}$ \bullet] Inserted ``that''
    \item[$4^{\text{th}}$ \bullet] Replaced $A(0,\varepsilon)$ 
    with $A(f,\varepsilon)$
    \item[$5^{\text{th}}$ \bullet] Matched the notation for the sample size to the notation for the corresponding algorithm before Section 3,  for example,
    $\widehat{A} \leftrightarrow \widehat{n}, \widetilde{A} \leftrightarrow \widetilde{n},
    A^* \leftrightarrow n^*$. As after section 3, we used a sequence $\bsn = \{n_0, n_1, \ldots \}.$ 
     \item[$6^{\text{th}}$ \bullet] The norm \emph{does} depend on the $\lambda_i$ because the $u_i$ depend on the $\lambda_i$.  This is clearer now that the norm for $\calF$ has been identified.
    \item[$7^{\text{th}}$ \bullet] Added $\exists \omega >0:$ to the definition
    \item[$8^{\text{th}}$ \bullet] Added for some $p>0$
    \item[$9^{\text{th}}$ \bullet] Changed ``no better than'' to ``no smaller than''
    \item[$10^{\text{th}}$ \bullet] Thanks for the correction. Changed to $1/2 -p$.
     \item[$11^{\text{th}}$ \bullet] Changed ``then'' to ``we have''
      \item[$12^{\text{th}}$ \bullet] Corrected the reference of the formula
      \item[$13^{\text{th}}$ \bullet] Regarding these proofs:
       \begin{itemize}
       \item[$-$] Changed the inequality to 
      $n + 2\le \textup{comp}(\calA(\calC),\varepsilon,\rho) + 2 \le n_j$ and corresponding places to correct the proof.
      \item[$-$] As $n$ is the sample size for a particular function of the optimal algorithm, it is possible 
      $n < \textup{cost}(A^*,\calC,\varepsilon,\rho) =  \textup{comp}(\calA(\calC),\varepsilon,\rho).$ 
           \item[$-$] Removed ``by (21)''
           \item[$-$] Added some derivation to illustrate it.
            \item[$-$] Changed to round brackets
            \item[$-$] Thanks for the suggestion. To make the proof clear, we used a different notation and way to present it.
       \end{itemize}
\end{enumerate}

@@

\subsection*{Report beginning  ``This article studies \ldots''}

Regarding the conditions (6) and (7), we have added some examples and explanation after we introduced these two conditions.

\begin{enumerate}
    \item Changed to ``Hence, the algorithm''
    \item Changed to ``and the algorithm''
    \item Changed to ``If for some $p > 0$ ''
    \item Thanks for the comment. Used a different function.
    \item Thanks for the comment. Added the examples.
    \item Have added the period at the end of the formula.
    \item Thanks for the comment. We added a clarification at the end of Section 2.
    \item Changed to $\lambda_i |\widehat{f}_i|$
    \item Changed the assumption to 
    $\lambda_i |\widehat{f}_i| =\Theta(i^{-p})$
    \item Removed $\forall j  \ge 1$
    \item Changed $\div$ to $/$
    \item There is no $+1$.  The derivation has been checked and expanded.
    \item Corrected the reference of the formula
    \item Erased $\le n_{j^\dagger}+1$ and clarified the argument leading to the inequality.
    \item Thanks for the suggestion. Changed $j^*$ to $j^{\ddagger}$
\end{enumerate}

\bigskip

\noindent With kind regards,\\[0.25cm]
The authors,

\noindent Y. Ding, F.J. Hickernell,  and Ll.A. Jim\'{e}nez Rugama.

\bibliographystyle{elsarticle-num}
\bibliography{FJH23,FJHown23}

### 2019-06-03 AutoAlgorithmLinearFunctionalDataVerF.tex
Updates from Overleaf

\title*{An Adaptive Algorithm Employing Continuous Linear Functionals}
\author{Yuhan Ding \and Fred J. Hickernell \and Llu\'{\i}s Antoni Jim\'{e}nez Rugama}
\institute{Yuhan Ding \at MCA 310, Department of Mathematics, Misericordia University,\\ 301 Lake St., Dallas, PA, 18612 \email{yding@misericordia.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 208, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and Llu\'{\i}s Antoni Jim\'{e}nez Rugama \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 208, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{ljimene1@hawk.iit.edu}}

@@

Our problem is not solvable for the whole Hilbert space, $\calF$, as can be demonstrated by contradiction. For any potential algorithm, we show that there exists some $f \in \calF$, that looks like $0$ to the algorithm, but for which $S(f)$ is far from $S(0) = 0$.   Choose any $A \in \calA(\calF)$ and $\varepsilon > 0$, and let $L_1, \ldots, L_n$ be the linear functionals that are used to compute $A(0,\varepsilon)$. Since the output space, $\calG$, is infinite dimensional and $n$ is finite, there exists some nonzero $f \in \calF$ satisfying that $L_1(f) = \cdots = L_n(f) = 0$ with non-zero $S(f)$. This means that $A(cf,\varepsilon) = A(0,\varepsilon)$ for any real $c$, and both of these have approximation error no greater than $\varepsilon$, i.e.,
\begin{align*}
    \varepsilon &\ge \frac 12 \left[ \DHJRnorm[\calG]{S(cf) - A(cf,\varepsilon)} + \DHJRnorm[\calG]{S(-cf) - A(-cf,\varepsilon)}\right] \\
    & =  \frac 12 \left[ \DHJRnorm[\calG]{cS(f) - A(0,\varepsilon)} + \DHJRnorm[\calG]{-cS(f) - A(0,\varepsilon)}\right] \\
    & \ge  \DHJRnorm[\calG]{cS(f)} = \DHJRabs{c}\DHJRnorm[\calG]{S(f)}  \qquad \text{by the triangle inequality}.
\end{align*}
Since $S(f) \ne 0$, it is impossible for this inequality to hold for all real $c$.  The presumed $A$ does not exist, $\calA(\calF)$ is empty, and our problem is not solvable for $\calF$. However, it is solvable for well-chosen subsets of $\calF$, as is shown in the sections below.

@@

Popular adaptive algorithms encountered typically employ heuristic error bounds.  While any algorithm can be fooled, we would like to have precise necessary conditions for being fooled, or equivalently, sufficient conditions for the algorithm to succeed.  Our adaptive algorithm has such conditions and follows in the vein of adaptive algorithms developed in \cite{HicEtal14a, HIcEtal14b, HicJim16a, JimHic16a}.

@@

\frednote{We need some horizontal axis labels for figures (d)--(f).} Figure \ref{fig:fourier_example} shows three functions and their coefficients for the case of Example \ref{exam:exam1} when $r=1$ and $n_j=2^j$. There are three functions: $f(x)={\rm e}^{-3x}\sin(3\pi x^2)$, $f_{\text{big}}=100f$ and $f_{\text{fuzzy}}$. The function $f_{\text{fuzzy}}$ is obtained by taking $\sigma_8(f_{\text{fuzzy}})=250\sigma_8(f)$ and $\sigma_j(f_{\text{fuzzy}})=\sigma_j(f)$ for $j \ne 8$.  Both $f$ and $f_{\text{big}}$ lie in the same cones, $\calC_{\bsn,a,b}$, and both appear similarly nice.  Therefore, we expect adaptive algorithms that are successful for $f$ to also be successful for $f_{\text{big}}$. However, $f_{\text{big}}$ does not lie in some of the balls,  $\calB_{\rho}$, that $f$ lies in.  On the other hand, the high frequency noise in $f_{\text{fuzzy}}$ suggests that only more robust and costly adaptive algorithms would be able to succeed for such an input.  This corresponds to the fact that $f_{\text{fuzzy}}$ does not lie in all cones, $\calC_{\bsn,a,b}$, that $f$ lies in.

@@

\begin{algo}\label{algo2}
Given $a$, $b$, the sequence $\bsn$, the cone $\calC$, the input function $f \in \calC$, and the absolute error tolerance $\varepsilon$, set $j=1$.
\begin{description}
\item[Step 1.] Compute $\sigma_{j}(f)$ as defined in \eqref{sumdef}.
\item[Step 2.] Check whether $j$ is large enough to satisfy the error tolerance, i.e.,
    \begin{equation*}\label{covcrit}
          \sigma_{j}(f) \le \frac{\varepsilon\sqrt{1 - b^2}}{ab} .
    \end{equation*}
    If this is true, then return $\widetilde{A}(f,\varepsilon) = A_{n_{j}}(f)$, where $A_n$ is defined in \eqref{optAdef}, and terminate the algorithm.
\item[Step 3.] Otherwise, increase $j$ by $1$ and return to Step $1$.
\end{description}
\end{algo}
\yuhannote{Please check the result of Theorem 1} 
\begin{theorem}\label{thm:compcost}
The algorithm, $\widetilde{A}$, defined in Algorithm \ref{algo2} lies in $\calA(\calC)$ and has computational cost $\textup{cost}(\widetilde{A},f,\varepsilon)=n_{j^*}$, where $j^*$ is defined implicitly by the inequalities 
\begin{equation} \label{eq:OurAlgjstar}
j^* = \min\left \{ j \in \bbN : \sigma_{j}(f) \le \frac{\varepsilon\sqrt{1 - b^2}}{ab}  \right\}.
\end{equation}
Moreover, $\textup{cost}(\widetilde{A},\calC,\varepsilon,\rho) \frednote{=} n_{j^\dagger}$, where $j^\dagger$ satisfies the following upper bound:
\begin{equation} \label{jdagger}
j^\dagger \le \min \left \{j \in \bbN : F(j) \ge \frac{\rho^2 a^4}{\varepsilon^2(1 - b^2)} \right\},
\end{equation}
% \begin{equation} \label{jdagger}
% j^\dagger \le \min \left \{j \in \bbN : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^4} F(j) \right\},
% \end{equation}
where $F$ is the strictly increasing function defined as
\begin{equation} \label{jsum}
F(j): = \sum_{k=0}^{j-1} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}, \qquad j \in \N.
\end{equation}
\end{theorem}

@@

The tighter upper bound in Theorem \ref{thm:compcost} may be obtained by a more careful argument in a similar vein.  
For any $f \in  \calC \cap \calB_\rho$ satisfying $n_{j^*} = \textup{cost}(\widetilde{A},f,\varepsilon) = \textup{cost}(\widetilde{A}, \calC, \varepsilon, \rho) = n_{j^{\dagger}}$ and for any $j < j^* = j^\dagger$,
\yuhannote{new proof starts here}\frednote{Looks good to me}
\begin{align*}
\rho^2 &\ge \DHJRnorm[\calF]{f}^2 = \DHJRbignorm[2]{\big( \widehat{f}_i \big)_{i=1}^{\infty}}^2  \ge \sum_{k=1}^j \DHJRbignorm[2]{\big( \widehat{f}_i \big)_{i=n_{k-1}+1}^{n_k}}^2 \\
& \ge \sum_{k=1}^j \frac{\sigma_k^2(f)}{\lambda_{n_{k-1}+1}^2} \qquad \text{by \eqref{normsigineq}}\\
& \ge \sum_{k=1}^{j-1} \frac{ b^{2(k-j)}\sigma_j^2(f)}{a^{2}\lambda_{n_{k-1}+1}^2} + \frac{\sigma_j^2(f)}{\lambda_{n_{j-1}+1}^2} \qquad \text{by \eqref{decayconedef}} \\
& = \sigma_j^2(f) \left[ \sum_{k=1}^{j-1} \frac{b^{2(k-j)}}{a^{2} \lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{j-1}+1}^2}\right]\\
&> \frac{ \sigma_j^2(f)}{a^2} \sum_{k=1}^{j} \frac{b^{2(k-j)}}{\lambda_{n_{k-1}+1}^2}  \qquad \text{since } a > 1\\
&= \frac{ \sigma_j^2(f)}{a^2}\sum_{k=0}^{j-1} \frac{b^{2(k+1-j)}}{\lambda_{n_{k}+1}^2}  \\
& \ge  \frac{b^2 \sigma_j^2(f)}{a^2} \sum_{k=0}^{j-1} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}  \qquad \text{since } \lambda_{n_k} \ge \lambda_{n_k+1}\\
& = \frac{b^2 \sigma_j^2(f)}{a^2} F(j)
\qquad \text{by } \eqref{jsum}.
\end{align*}

@@

\yuhannote{Also change the analysis and corollary 1}
The upper bound on $\textup{cost}(\widetilde{A},\calC,\rho,\varepsilon)$ in Theorem \ref{thm:compcost}  is a non-decreasing function of $\rho/\varepsilon$, which depends on the behavior of the sequence $(\lambda_{n_j})_{j=0}^\infty$.  This in turn depends both on the increasing sequence $\bsn$ and on the non-increasing sequence $(\lambda_i)_{i=1}^\infty$. Considering the sum $F(j)$,
one can imagine that in some cases the first term in the sum $F(j)$ dominates, while in other cases the last term in the sum dominates, all depending on how $b^{k-j}/\lambda_{n_{k}}$ behaves with $k$ and $j$.  These simplifications lead to two simpler, but coarser upper bounds on the cost of $\widetilde{A}$.

@@

\begin{proof}
Ignoring all but the first term in the definition of $F$ in \eqref{jsum} implies that 
\begin{equation*} \label{jsumupperbd}
j^\dagger \le \min \left \{j \in \bbN : b^{-2j}  \ge  \frac{\rho^2 a^4 \lambda_{n_{0}}^2 }{\varepsilon^2(1 - b^2)} \right\}.
\end{equation*}
This implies \eqref{jdaggerfirst}.

Ignoring all but the last term of the sum \eqref{jsum}. leads to the simpler upper bound similar to  \eqref{jdaggerlast}
\begin{equation*} 
j^\dagger \le \min \left \{j \in \bbN :  \lambda_{n_{j-1}}^{-1} \ge \frac{ \rho a^2b}{ \varepsilon\sqrt{1 - b^2}} \right\}.
\end{equation*}
If the $\lambda_{n_{j-1}}$ decay as assumed in \eqref{lambdankbd} then
\[
j^\dagger \le \min \left \{j \in \bbN :  \alpha \beta^{-j} \le \frac{\rho a^2b }{ \varepsilon\sqrt{1 - b^2}} \right\},
\]
which implies \eqref{jdaggerlastsimple}.
\end{proof}

@@

\yuhannote{corollary 2 is not changed} \frednote{I think we need a small change}
\begin{corollary} \label{cor:tAsameCosthA} Suppose that the sequence $\bsn$ is chosen to satisfy
\begin{equation} \label{lambdaDecay}
\lambda_{n_{j+1}} \ge c_\lambda \lambda_{n_j}, \qquad j \in \bbN, 
\end{equation}
for some positive $c_\lambda$.  Then $\textup{cost}(\widetilde{A},\calC, \varepsilon,\rho)$ is essentially no worse than \linebreak[4]
$\textup{cost}(\widehat{A},\calB_{\rho},\varepsilon)$ in the sense of \eqref{essNoWorseDef}. 
\end{corollary}

\begin{proof}
Combining the upper bound on $n_{j^\dagger} = \textup{cost}(\widetilde{A},\calC,\varepsilon,\rho)$ in \eqref{jdaggerlast} plus  \eqref{lambdaDecay} above, it follows that
\begin{equation*}
\lambda_{n_{j^\dagger{{\color{red}+1}}}} \ge c_{\lambda}^{{\color{red}{3}}} \lambda_{n_{j^\dagger-2}} > \frac{\varepsilon c_{\lambda}^{{\color{red}{3}}}\sqrt{1 - b^2}}{\rho ab } = \frac{\omega \varepsilon}{\rho} \ge \lambda_{\widehat{n}+1},
\qquad \omega := \frac{ c_{\lambda}^{{\color{red}{3}}}\sqrt{1 - b^2}}{ab }
\end{equation*}
where $\widehat{n}
= \textup{cost}(\widehat{A},\calB_{\rho},\omega\varepsilon )$.
Since the $\lambda_i$ are non-increasing, {\color{red} $n_{j^\dagger+1} \le \widehat{n}+1$, and so $n_{j^\dagger} \le \widehat{n}$.  Thus, }
\begin{equation*}
\textup{cost}(\widetilde{A},\calC,\varepsilon,\rho) = n_{j^\dagger} \le \widehat{n} = \textup{cost}(\widehat{A},\calB_{\rho}, \omega \varepsilon).
\end{equation*}
\end{proof}

@@

A lower bound on $\textup{comp}(\calA(\calC),\varepsilon,\rho)$ is established by constructing fooling functions in $\calC$ with norms no greater than $\rho$.  To obtain a result that can be compared with the cost of our algorithm, we assume that 
\begin{equation} \label{lambdaRatio}
R = \sup_{j \in \bbN} \frac{\lambda_{n_{j-1}}}{\lambda_{n_j}} < \infty.
\end{equation}
\frednote{That is, the subsequence $\lambda_{n_1}, \lambda_{n_2}, \ldots$ cannot decay too quickly.}

@@

\yuhannote{Please check the definition of $j^\ddagger$}
\begin{theorem} \label{thm:lowbdcomp}
Under assumption  \eqref{lambdaRatio}, a lower bound on the complexity of the linear problem defined in \eqref{probDef} is
\begin{align*}
 %\label{compbdA}
&\textup{comp}(\calA(\calC),\varepsilon,\rho) \ge n_{j^{\ddagger}} - 1, 
\intertext{where}
%\label{compbdB}
j^{\ddagger} & = \min \left \{ j \in \bbN : F(j+2) \ge \frac{\rho^2}{\varepsilon^2 b^2}
\left[\frac{(a+1)^{2} R^2 }{(a-1)^2} + 1\right]^{-1}
\right \},
\end{align*}
where $F(j)$ defined in \eqref{jsum}.
\end{theorem}

@@

\begin{acknowledgement}
The authors would like to thank three anonymous referees for their suggestions which helped to improve the presentation of the results.

>>>>

\title*{An Adaptive Algorithm Employing Continuous Linear Functionals}
\author{Yuhan Ding \and Fred J. Hickernell \and Llu\'{\i}s Antoni Jim\'{e}nez Rugama}
\institute{Yuhan Ding \at Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 208, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{yding2@hawk.iit.edu}
\and
Fred J. Hickernell \at Center for Interdisciplinary Scientific Computation and \\
Department of Applied Mathematics, Illinois Institute of Technology \\ RE 208, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{hickernell@iit.edu}
\and Llu\'{\i}s Antoni Jim\'{e}nez Rugama \at
Department of Applied Mathematics, Illinois Institute of Technology,\\ RE 208, 10 W.\ 32$^{\text{nd}}$ St., Chicago, IL 60616 \email{ljimene1@hawk.iit.edu}}

@@

Our problem is not solvable for the whole Hilbert space, $\calF$, as can be demonstrated by contradiction. For any potential algorithm, we show that there exists some $f \in \calF$, that looks like $0$ to the algorithm, but for which $S(f)$ is far from $S(0) = 0$.   Choose any $A \in \calA(\calF)$ and $\varepsilon > 0$, and let $L_1, \ldots, L_n$ be the linear functionals that are used to compute $A(0,\varepsilon)$. Since the output space, $\calG$, is infinite dimensional and $n$ is finite, there exists some nonzero $f \in \calF$ satisfying that $L_1(f) = \cdots = L_n(f) = 0$ with non-zero $S(f)$. This means that $A(\pm cf,\varepsilon) = A(0,\varepsilon)$ for any real $c$, and $A(\pm cf,\varepsilon)$ both of have approximation error no greater than $\varepsilon$, i.e.,
\begin{align*}
    \varepsilon &\ge \frac 12 \left[ \DHJRnorm[\calG]{S(cf) - A(cf,\varepsilon)} + \DHJRnorm[\calG]{S(-cf) - A(-cf,\varepsilon)}\right] \\
    & =  \frac 12 \left[ \DHJRnorm[\calG]{cS(f) - A(0,\varepsilon)} + \DHJRnorm[\calG]{-cS(f) - A(0,\varepsilon)}\right] \\
    & \ge  \DHJRnorm[\calG]{cS(f)} = \DHJRabs{c}\DHJRnorm[\calG]{S(f)}  \qquad \text{by the triangle inequality}.
\end{align*}
Since $S(f) \ne 0$, it is impossible for this inequality to hold for all real $c$.  The presumed $A$ does not exist, $\calA(\calF)$ is empty, and our problem is not solvable for $\calF$. However, our problem is solvable for well-chosen subsets of $\calF$, as is shown in the sections below.

@@

Popular adaptive algorithms encountered typically employ heuristic error bounds.  While any algorithm can be fooled, we would like to have precise necessary conditions for being fooled, or equivalently, sufficient conditions for the algorithm to succeed.  Our adaptive algorithm has such conditions and follows in the vein of adaptive algorithms developed in \cite{HicEtal14a, HIcEtal14b, HicJim16a, JimHic16a, HicEtal17a}.

@@

Figure \ref{fig:fourier_example} shows three functions and their coefficients for the case of Example \ref{exam:exam1} when $r=1$ and $n_j=2^j$. These  functions are $f(x)={\rm e}^{-3x}\sin(3\pi x^2)$, $f_{\text{big}}=100f$ and $f_{\text{fuzzy}}$. The function $f_{\text{fuzzy}}$ is obtained by taking $\sigma_8(f_{\text{fuzzy}})=250\sigma_8(f)$ and $\sigma_j(f_{\text{fuzzy}})=\sigma_j(f)$ for $j \ne 8$.  Both $f$ and $f_{\text{big}}$ lie in the same cones, $\calC_{\bsn,a,b}$, for all $\bsn$, $a$, and $b$, and both appear similarly nice to the eye.  Therefore, we expect adaptive algorithms that are successful for $f$ to also be successful for $f_{\text{big}}$. However, $f_{\text{big}}$ does not lie in some of the balls,  $\calB_{\rho}$, that $f$ lies in.  On the other hand, the high frequency noise in $f_{\text{fuzzy}}$ suggests that only more robust and costly adaptive algorithms would succeed for such an input.  This corresponds to the fact that $f_{\text{fuzzy}}$ does not lie in all cones, $\calC_{\bsn,a,b}$, that $f$ lies in.

@@

\begin{algo}\label{algo2}
Given $a$, $b$, the sequence $\bsn$, the cone $\calC$, the input function $f \in \calC$, and the absolute error tolerance $\varepsilon$, set $j=1$.
\begin{description}
\item[Step 1.] Compute $\sigma_{j}(f)$ as defined in \eqref{sumdef}.
\item[Step 2.] Check whether $j$ is large enough to satisfy the error tolerance, i.e.,
    \begin{equation*}\label{covcrit}
          \sigma_{j}(f) \le \frac{\varepsilon\sqrt{1 - b^2}}{ab} .
    \end{equation*}
    If this is true, then return $\widetilde{A}(f,\varepsilon) = A_{n_{j}}(f)$, where $A_n$ is defined in \eqref{optAdef}, and terminate the algorithm.
\item[Step 3.] Otherwise, increase $j$ by $1$ and return to Step $1$.
\end{description}
\end{algo}

\begin{theorem}\label{thm:compcost}
The algorithm, $\widetilde{A}$, defined in Algorithm \ref{algo2} lies in $\calA(\calC)$ and has computational cost $\textup{cost}(\widetilde{A},f,\varepsilon)=n_{j^*}$, where $j^*$ is defined implicitly by the inequalities 
\begin{equation} \label{eq:OurAlgjstar}
j^* = \min\left \{ j \in \bbN : \sigma_{j}(f) \le \frac{\varepsilon\sqrt{1 - b^2}}{ab}  \right\}.
\end{equation}
Moreover, $\textup{cost}(\widetilde{A},\calC,\varepsilon,\rho) = n_{j^\dagger}$, where $j^\dagger$ satisfies the following upper bound:
\begin{equation} \label{jdagger}
j^\dagger \le \min \left \{j \in \bbN : F(j) \ge \frac{\rho^2 a^4}{\varepsilon^2(1 - b^2)} \right\},
\end{equation}
% \begin{equation} \label{jdagger}
% j^\dagger \le \min \left \{j \in \bbN : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^4} F(j) \right\},
% \end{equation}
where $F$ is the strictly increasing function defined as
\begin{equation} \label{jsum}
F(j): = \sum_{k=0}^{j-1} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}, \qquad j \in \N.
\end{equation}
\end{theorem}

@@

The tighter upper bound in Theorem \ref{thm:compcost} may be obtained by a more careful argument in a similar vein.  
For any $f \in  \calC \cap \calB_\rho$ satisfying $n_{j^*} = \textup{cost}(\widetilde{A},f,\varepsilon) = \textup{cost}(\widetilde{A}, \calC, \varepsilon, \rho) = n_{j^{\dagger}}$ and for any $j < j^* = j^\dagger$,
\begin{align*}
\rho^2 &\ge \DHJRnorm[\calF]{f}^2 = \DHJRbignorm[2]{\big( \widehat{f}_i \big)_{i=1}^{\infty}}^2  \ge \sum_{k=1}^j \DHJRbignorm[2]{\big( \widehat{f}_i \big)_{i=n_{k-1}+1}^{n_k}}^2 \\
& \ge \sum_{k=1}^j \frac{\sigma_k^2(f)}{\lambda_{n_{k-1}+1}^2} \qquad \text{by \eqref{normsigineq}}\\
& \ge \sum_{k=1}^{j-1} \frac{ b^{2(k-j)}\sigma_j^2(f)}{a^{2}\lambda_{n_{k-1}+1}^2} + \frac{\sigma_j^2(f)}{\lambda_{n_{j-1}+1}^2} \qquad \text{by \eqref{decayconedef}} \\
& = \sigma_j^2(f) \left[ \sum_{k=1}^{j-1} \frac{b^{2(k-j)}}{a^{2} \lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{j-1}+1}^2}\right]\\
&> \frac{ \sigma_j^2(f)}{a^2} \sum_{k=1}^{j} \frac{b^{2(k-j)}}{\lambda_{n_{k-1}+1}^2}  \qquad \text{since } a > 1\\
&= \frac{ \sigma_j^2(f)}{a^2}\sum_{k=0}^{j-1} \frac{b^{2(k+1-j)}}{\lambda_{n_{k}+1}^2}  \\
& \ge  \frac{b^2 \sigma_j^2(f)}{a^2} \sum_{k=0}^{j-1} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}  \qquad \text{since } \lambda_{n_k} \ge \lambda_{n_k+1}\\
& = \frac{b^2 \sigma_j^2(f)}{a^2} F(j)
\qquad \text{by } \eqref{jsum}.
\end{align*}

@@

The upper bound on $\textup{cost}(\widetilde{A},\calC,\rho,\varepsilon)$ in Theorem \ref{thm:compcost}  is a non-decreasing function of $\rho/\varepsilon$, which depends on the behavior of the sequence $(\lambda_{n_j})_{j=0}^\infty$.  This in turn depends both on the increasing sequence $\bsn$ and on the non-increasing sequence $(\lambda_i)_{i=1}^\infty$. Considering the sum $F(j)$,
one can imagine that in some cases the first term in the sum $F(j)$ dominates, while in other cases the last term in the sum dominates, all depending on how $b^{k-j}/\lambda_{n_{k}}$ behaves with $k$ and $j$.  These simplifications lead to two simpler, but coarser upper bounds on the cost of $\widetilde{A}$.

@@

\begin{proof}
Ignoring all but the first term in the definition of $F$ in \eqref{jsum} implies that 
\begin{align*} \label{jsumupperbd}
j^\dagger & \le \min \left \{j \in \bbN : b^{-2j}  \ge  \frac{\rho^2 a^4 \lambda_{n_{0}}^2 }{\varepsilon^2(1 - b^2)} \right\} \\
& = 
\min \left \{j \in \bbN : j  \ge  \log\left(\frac{\rho a^2 \lambda_{n_{0}} }{\varepsilon(\sqrt{1 - b^2}}\right) \middle/ \log\left(\frac 1b \right) \right\}.
\end{align*}
This implies \eqref{jdaggerfirst}.

Ignoring all but the last term of the sum \eqref{jsum}. leads to the simpler upper bound similar to  \eqref{jdaggerlast}: 
\begin{equation*} 
j^\dagger \le \min \left \{j \in \bbN :  \lambda_{n_{j-1}}^{-1} \ge \frac{ \rho a^2b}{ \varepsilon\sqrt{1 - b^2}} \right\}.
\end{equation*}
If the $\lambda_{n_{j-1}}$ decay as assumed in \eqref{lambdankbd} then
\[
j^\dagger \le \min \left \{j \in \bbN :  \alpha \beta^{-j} \le \frac{\rho a^2b }{ \varepsilon\sqrt{1 - b^2}} \right\},
\]
which implies \eqref{jdaggerlastsimple}.
\end{proof}

@@

\begin{corollary} \label{cor:tAsameCosthA} Suppose that the sequence $\bsn$ is chosen to satisfy
\begin{equation} \label{lambdaDecay}
\lambda_{n_{j+1}} \ge c_\lambda \lambda_{n_j}, \qquad j \in \bbN, 
\end{equation}
for some positive $c_\lambda$.  Then $\textup{cost}(\widetilde{A},\calC, \varepsilon,\rho)$ is essentially no worse than \linebreak[4]
$\textup{cost}(\widehat{A},\calB_{\rho},\varepsilon)$ in the sense of \eqref{essNoWorseDef}. 
\end{corollary}

\begin{proof}
Combining the upper bound on $n_{j^\dagger} = \textup{cost}(\widetilde{A},\calC,\varepsilon,\rho)$ in \eqref{jdaggerlast} plus  \eqref{lambdaDecay} above, it follows that
\begin{equation*}
\lambda_{n_{j^\dagger}} \ge c_{\lambda}^{2} \lambda_{n_{j^\dagger-2}} > \frac{\varepsilon c_{\lambda}^{2}\sqrt{1 - b^2}}{\rho ab } = \frac{\omega \varepsilon}{\rho} \ge \lambda_{\widehat{n}+1},
\qquad \omega := \frac{ c_{\lambda}^{2}\sqrt{1 - b^2}}{ab },
\end{equation*}
where $\widehat{n}
= \textup{cost}(\widehat{A},\calB_{\rho},\omega\varepsilon )$.
Since the $\lambda_i$ are non-increasing and $\lambda_{n_{j^\dagger}} > \lambda_{\hat{n}+1}$, it follows that  $n_{j^\dagger} < \widehat{n}+1$, and so $n_{j^\dagger} \le \widehat{n}$.  Thus, 
\begin{equation*}
\textup{cost}(\widetilde{A},\calC,\varepsilon,\rho) = n_{j^\dagger} \le \widehat{n} = \textup{cost}(\widehat{A},\calB_{\rho}, \omega \varepsilon).
\end{equation*}
\end{proof}

@@

A lower bound on $\textup{comp}(\calA(\calC),\varepsilon,\rho)$ is established by constructing fooling functions in $\calC$ with norms no greater than $\rho$.  To obtain a result that can be compared with the cost of our algorithm, we assume that 
\begin{equation} \label{lambdaRatio}
R = \sup_{j \in \bbN} \frac{\lambda_{n_{j-1}}}{\lambda_{n_j}} < \infty.
\end{equation}
That is, the subsequence $\lambda_{n_1}, \lambda_{n_2}, \ldots$ cannot decay too quickly.

@@

\begin{theorem} \label{thm:lowbdcomp}
Under assumption  \eqref{lambdaRatio}, a lower bound on the complexity of the linear problem defined in \eqref{probDef} is
\begin{align*}
 %\label{compbdA}
&\textup{comp}(\calA(\calC),\varepsilon,\rho) \ge n_{j^{\ddagger}} - 1, 
\intertext{where}
%\label{compbdB}
j^{\ddagger} & = \min \left \{ j \in \bbN : F(j+2) \ge \frac{\rho^2}{\varepsilon^2 b^2}
\left[\frac{(a+1)^{2} R^2 }{(a-1)^2} + 1\right]^{-1}
\right \},
\end{align*}
and $F(j)$ is defined in \eqref{jsum}.
\end{theorem}

@@

\begin{acknowledgement}
The authors would like to thank the organizers for a wonderful MCQMC 2018. 
We also thank the referees for their many helpful suggestions.  This work is supported in part by National Science Foundation grants DMS-1522687 and DMS-1638521 (SAMSI).  Yuhan Ding acknowledges the support of the Department of Mathematics, Misercordia University.
